Cheminformatics: The Computer Science of Chemical Discovery

### 2011-11-09 rebuttal.tex

\documentclass[11pt,a4paper]{article}
\usepackage{times}

\setlength{\parindent}{0cm}

\begin{document}
\newcommand{\comment}[1]{\textit{``\ldots #1''}\par\vspace{0.5em}}
\newcommand{\response}[1]{#1\vspace{1em}}

\fbox{\textbf{Reviewer 1}}
\
\comment{when the authors state 'together with an even larger number
  of annotations', they shold explain what are annotations and how
  much is the large number;} 
\response{We have updated the text on page 1 to be more explicit in
  terms of what annotations are and provide a concete example of the
  sizes involved.}

@@

\comment{For an overview I would expect a more comprehensive set of
  references.}
\response{We agree with the reviewer that a more comprehensive list of
references would be appropriate and useful. However, due to format
limitations of the article, we have been restricted to the use of XXX
references. To alleviate this problem, we have included a supplement
that provides a more extensive list of references, specifically
including a number of texts covering cheminformatics broadly.}

\fbox{\textbf{Reviewer 2}}

\comment{The authors mention limitations of SMILES implementations,
  but do not mention canonicalization in this context, often using the
  Morgan algorithm published in 1965 in J. Chem. Doc.}
\response{We have updated the text on page 3 to note the Morgan
  algorithm in the context of chemical structure representation}

\comment{The authors state that 3D information is lost when only
  considering the molecular graph. However, they do not make it clear
  that the 3D space is implicitly encoded in the molecular graph –
  this is essential to cover properly.}
\response{We have update the text on page 2 to note that 3D
  information is implicit in 2D representations}

### 2011-11-09 rebuttal.tex

\fbox{\textbf{Reviewer 2}}

\comment{The authors mention limitations of SMILES implementations,
  but do not mention canonicalization in this context, often using the
  Morgan algorithm published in 1965 in J. Chem. Doc.}
\response{We have updated the text on page 3 to note the Morgan
  algorithm in the context of chemical structure representation}

\comment{The authors state that 3D information is lost when only
  considering the molecular graph. However, they do not make it clear
  that the 3D space is implicitly encoded in the molecular graph –
  this is essential to cover properly.}
\response{We have update the text on page 2 to note that 3D
  information is implicit in 2D representations}

\comment{If this manuscript were to be published it would be
  completely rewriting the history of cheminformatics}
\response{While we agree with the reviewer that some key milestones in
  the history of cheminformatics were not included in the original
  article, we believe that the addition of the supplemental history
  addresses this problem. Furthermore, our original aim was \emph{not}
  to provide a complete historical overview of the field. Others have
  already provided such overviews. Our aim was to present a brief
  overview of \emph{topics} in cheminformatics, selected because we
  believed that they represent on going challenges for the field and
  thus amenable to coss-disciplinary efforts.}

### 2011-11-09 rebuttal.tex

\comment{The authors state that 3D information is lost when only
  considering the molecular graph. However, they do not make it clear
  that the 3D space is implicitly encoded in the molecular graph –
  this is essential to cover properly.}
\response{We have update the text on page 2 to note that 3D
  information is implicit in 2D representations}

\comment{If this manuscript were to be published it would be
  completely rewriting the history of cheminformatics}
\response{While we agree with the reviewer that some key milestones in
  the history of cheminformatics were not included in the original
  article, we believe that the addition of the supplemental history
  addresses this problem. Furthermore, our original aim was \emph{not}
  to provide a complete historical overview of the field. Others have
  already provided such overviews. Our aim was to present a brief
  overview of \emph{topics} in cheminformatics, selected because we
  believed that they represent on going challenges for the field and
  thus amenable to coss-disciplinary efforts.}

\fbox{\textbf{Reviewer 3}}

\comment{page 1, column 1, line 24: the authors need to provide a
  reference for the assertion that cheminformatics is an older field,
  particularly given the comments later in this
  manuscript.}
\response{We have removed the assertion that cheminformatics is older
  than bioinformatics. Primarily, because there is no obvious
  reference that claims this. But alse because, depending on what one
  considers bioinformatics and cheminformatics, one field can be
  identified as being older or younger than the other. We feel that
  claims on age do not strengthen the paper and have thus removed it}

\comment{p1, c1, l29-37: I would dispute that until recently the
  cheminformatics techniques have been closely guarded
  secrets. J. Chem. Doc. was founded in 1961 and other techniques even
  older, such as the Wiener index (1947), Wiswesser Line Notation
  (1949), etc. Not to mention the actual foundations of what we call
  cheminformatics back to the atomistic theory of the 19th century.}

### 2011-11-09 paper.tex
more rebuttal

\maketitle
\begin{abstract}
  One of the most prominent success stories in all the sciences over
  the last decade has been the advance of bioinformatics: The
  interdisciplinary collaboration between computer scientists and
  molecular biologists led as one of its major accomplishments to the
  sequencing of the human genome, where researchers connected
  biological concepts to the theory of string algorithms. However,
  despite this great success, few computer scientists are familiar
  with a related (and older!)  discipline: cheminformatics, the use of
  computers to represent the structures of small molecules and analyze
  their properties.  Cheminformatics has wide applicability, from the
  discovery of novel drug candidates to the optimization of certain
  physicochemical properties of small molecules. Until recently, the
  data and techniques employed in the cheminformatics field have been
  closely guarded secrets of companies whose financial success
  depended on being the first to produce the new ``miracle molecule.''
  Only within the last decade---and as an effect of a change in
  mindsets, government-mandated policy changes and last but not least
  because of chemists volunteering their time for an Open Science
  ``movement''---have researchers gained access to freely available
  software packages and databases of tens of millions of
  chemicals. Academic chemists now confront a variety of unsolved
  algorithmic problems that could not have been tackled a decade ago,
  but whose solutions are critical to research ranging from
  determining the behavior of small molecules in biological pathways,
  to finding a cure for malaria and other ``neglected'' diseases.
\end{abstract}

>>>>

\maketitle
\begin{abstract}
  One of the most prominent success stories in all the sciences over
  the last decade has been the advance of bioinformatics: The
  interdisciplinary collaboration between computer scientists and
  molecular biologists led as one of its major accomplishments to the
  sequencing of the human genome, where researchers connected
  biological concepts to the theory of string algorithms. However,
  despite this great success, few computer scientists are familiar
  with a related discipline: cheminformatics, the use of computers to
  represent the structures of small molecules and analyze their
  properties.  Cheminformatics has wide applicability, from the
  discovery of novel drug candidates to the optimization of certain
  physicochemical properties of small molecules. Until recently, the
  data and techniques employed in the cheminformatics field have been
  closely guarded secrets of companies whose financial success
  depended on being the first to produce the new ``miracle molecule.''
  Only within the last decade---and as an effect of a change in
  mindsets, government-mandated policy changes and last but not least
  because of chemists volunteering their time for an Open Science
  ``movement''---have researchers gained access to freely available
  software packages and databases of tens of millions of
  chemicals. Academic chemists now confront a variety of unsolved
  algorithmic problems that could not have been tackled a decade ago,
  but whose solutions are critical to research ranging from
  determining the behavior of small molecules in biological pathways,
  to finding a cure for malaria and other ``neglected'' diseases.
\end{abstract}
